---
title: "p8105_hw2_lc4152"
output: github_document
---

## Problem 1

```{r}
library(tidyverse)
library(readxl)
```

```{r}
subway_df = 
  read_csv(file = "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
  select(line:entry, vending, ada) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
dim(subway_df)
```
This dataset contains information about NYC subway entrances and exits, including the line, station name, station coordinates, routes served, entry permission, vending availability, entrance type, and ADA compliance. I cleaned the column names, selected the required variables, and converted entry from YES/NO to TRUE/FALSE. The resulting dataset has 1,868 rows and 19 columns. The data are not yet tidy because routes served are spread across multiple columns, from route1 to route11.

Number of stations:
```{r}
subway_df |>
  distinct(line, station_name) |>
  nrow()
```

Number of stations are ADA compliant:
```{r}
subway_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

Proportion:
```{r}
subway_df |>
  filter(vending == "NO") |>
  summarize(
    proportion_allow_entry = mean(entry)
  )
```

Reformat route:
```{r}
subway_long_df = 
  subway_df |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name",
    values_transform = list(route_name = as.character),
    values_drop_na = TRUE
  )
```

Number of distinct stations serve A train:
```{r}
subway_long_df |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()
```

Number of ADA compliant among those serve A:
```{r}
subway_long_df |>
  filter(route_name == "A", ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

There are 465 distinct stations, of which 84 are ADA compliant. Among entrances and exits without vending, 37.7% allow entry. After converting the route columns to long format, I found that 60 distinct stations serve the A train, and 17 of these stations are ADA compliant.

## Problem 2

```{r}
mr_trash_df = 
  read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N"),
    skip = 1)

```






















## Problem 3