---
title: "p8105_hw2_lc4152"
output: github_document
---

## Problem 1

```{r}
library(tidyverse)
library(readxl)
```

```{r}
subway_df = 
  read_csv(file = "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
  select(line:entry, vending, ada) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
dim(subway_df)
```
This dataset contains information about NYC subway entrances and exits, including the line, station name, station coordinates, routes served, entry permission, vending availability, entrance type, and ADA compliance. I cleaned the column names, selected the required variables, and converted entry from YES/NO to TRUE/FALSE. The resulting dataset has 1,868 rows and 19 columns. The data are not yet tidy because routes served are spread across multiple columns, from route1 to route11.

Number of stations:
```{r}
subway_df |>
  distinct(line, station_name) |>
  nrow()
```

Number of stations are ADA compliant:
```{r}
subway_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

Proportion:
```{r}
subway_df |>
  filter(vending == "NO") |>
  summarize(
    proportion_allow_entry = mean(entry)
  )
```

Reformat route:
```{r}
subway_long_df = 
  subway_df |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name",
    values_transform = list(route_name = as.character),
    values_drop_na = TRUE
  )
```

Number of distinct stations serve A train:
```{r}
subway_long_df |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()
```

Number of ADA compliant among those serve A:
```{r}
subway_long_df |>
  filter(route_name == "A", ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

There are 465 distinct stations, of which 84 are ADA compliant. Among entrances and exits without vending, 37.7% allow entry. After converting the route columns to long format, I found that 60 distinct stations serve the A train, and 17 of these stations are ADA compliant.

## Problem 2

Mr.Trash:
```{r}
mr_trash_df = 
  read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N"),
    skip = 1
  ) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.integer(year),
    month = str_trim(month),
    sports_balls = as.integer(round(sports_balls)),
    trash_wheel = "Mr. Trash Wheel"
  )

```

Professor trash:
```{r}
professor_trash_df =
  read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = cell_cols("A:M"),
    skip = 1
  ) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.integer(year),
    month = str_trim(month),
    plastic_bottles = as.numeric(plastic_bottles),
    trash_wheel = "Professor Trash Wheel"
  )
```

Gwynnda:
```{r}
gwynnda_df =
  read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynnda the Good Wheel of the W",
    range = cell_cols("A:L"),
    skip = 1
  ) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster)) |>
  mutate(
    year = as.integer(year),
    month = str_trim(month),
    trash_wheel = "Gwynnda"
  )
```

Combine:
```{r}
trash_wheel_df =
  bind_rows(
    mr_trash_df,
    professor_trash_df,
    gwynnda_df
  )

dim(trash_wheel_df)
```

```{r}
professor_weight =
  trash_wheel_df |>
  filter(trash_wheel == "Professor Trash Wheel") |>
  summarize(
    total_weight = sum(weight_tons, na.rm = TRUE)
  ) |>
  pull(total_weight)

professor_weight
```

```{r}
gwynnda_cigarettes =
  trash_wheel_df |>
  filter(
    trash_wheel == "Gwynnda",
    year == 2022,
    month == "June"
  ) |>
  summarize(
    total_cigarettes = sum(cigarette_butts, na.rm = TRUE)
  ) |>
  pull(total_cigarettes)

gwynnda_cigarettes
```

The combined dataset contains `r nrow(trash_wheel_df)` observations from Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda. Each observation represents one dumpster collection. Key variables include the collection date, weight in tons, volume in cubic yards, and counts of items such as plastic bottles and cigarette butts. I cleaned the column names, removed rows without dumpster-specific data, standardized the year and month variables, and rounded sports balls to integers. I also added a variable identifying each trash wheel before combining the datasets. Professor Trash Wheel collected a total of `r professor_weight` tons of trash, and Gwynnda collected `r gwynnda_cigarettes` cigarette butts in June 2022.


## Problem 3






















